---
title: Bike Share in Toronto
author: Mark Edney
date: '2022-02-19'
slug: []
categories:
  - Project
tags:
  - R
  - Shiny App
draft: yes
description: ''
archives:
  - 2022/02
---
> This article is based on a project written on 01/14/2021

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning = FALSE)
```

## Bike Rental Shiny App

This application use the data collected from the [Toronto Open Data](https://open.toronto.ca/dataset/bike-share-toronto-ridership-data/) to 
generate a histogram of the usage of rental bikes in Toronto during the month of 
June in 2020.
```{r packages, results='hide'}
install.packages("opendatatoronto", 
                 repos = "http://cran.us.r-project.org",
                 dependencies = TRUE)
library(opendatatoronto)
library(tidyverse)
library(lubridate)
library(shiny)
```
## UI

There are two user inputs on the UI side:

- A slider that limits the maximum and minimum of the displayed values
- A checkbox that excludes users with a annual bike pass

```{r UI, eval=FALSE}
        sidebarPanel(
            sliderInput("dur",
                        "Trip Duration:",
                        min = 0,
                        max = 500,
                        value = c(0,500)),
            checkboxInput("freq",
                        "Exclude annual users:",
                        value = FALSE))
```

## Server

```{r Serverhide, echo = FALSE}
 # get package
    package <- show_package("7e876c24-177c-4605-9cef-e50dd74c617f")
    
    # get all resources for this package
    resources <- list_package_resources("7e876c24-177c-4605-9cef-e50dd74c617f")
    # identify datastore resources; by default, Toronto Open Data sets datastore resource format to CSV for non-geospatial and GeoJSON for geospatial resources
    datastore_resources <- filter(resources, tolower(format) %in% c('zip', 'geojson'))
    # load the first datastore resource as a sample
    data <- filter(datastore_resources, name == "Bike share ridership 2020") %>% get_resource()
    data2 <-  data$`2020-06.csv`
    data2[grepl("Time",names(data2))] <- 
        lapply(data2[grepl("Time",names(data2))], parse_date_time, orders = "mdy HM")
    data2$Dur <- as.numeric(data2$End.Time - data2$Start.Time,units="mins")

```

```{r app, echo=FALSE, warning=FALSE}
knitr::include_url("https://m2edney.shinyapps.io/Toronto_bike/?_ga=2.75026646.519818068.1645317226-1341333380.1645206372")
```