---
title: Benchmarking Data Tables
author: Mark Edney
date: '2022-04-12'
slug: []
categories:
  - How-to
tags:
  - R
draft: yes
description: 'A look into the claim performance of Data Tables in the R language.'
archives:
  - 2022/04
---
When I started learning R, I was heard vague tales of the use of Data tables. Really just whisperers of something to consider in the future after I've become more proficient. Well now is the time to learn what if anything I've been missing out on.

## Introduction

Data tables are a potential replacement for the common dataframe. It  seeks to perform that same role but with improved performance. I would like to see the speed comparison between data frames, data tables and tibbles. I will use the `microbenchmark` package to perform the actual benchmarking.

```{r packages, warning=FALSE, message=FALSE}
library(tidyverse)
library(data.table)
library(microbenchmark)
library(farff)
```

For the benchmark, I will use the 'credit-g' dataset which can be found on the [open ml](https://www.openml.org/search?type=data&status=active&id=31) website. I'm pretty sure the last openml dataset I used was a csv file but the seem to have moved to a ARFF format. I will need to use the `farff` package to load the data. 

```{r load, message=FALSE}
df <- farff::readARFF('dataset_31_credit-g.arff')
dt <- setDT(df)
ti <- tibble(df)
```

## Syntax

The syntax for data tables is a little be different

> DT[i,j,by]

In this manner, a at data table can be subset by i, to calculate j when grouped with a by. For this comparison, we will look at the performance of finding the average age of the credit holders grouped by the class or credit rating.

Along with the special syntax, there are some common functions that add some additional simplification.

> .()

The '.()' function can be used as a place holder for 'list()'. The list function is usefull for subseting


## Grouped Aggregate

Aggregating data in data tables is simple by using the j and by parameters in the syntax. Again multiple functions or even multiple groupings can be passed with the '.()' function 

```{r grouped}
microbenchmark(Data_Frame = df %>% 
                                 group_by(class) %>%
                       summarise(avg = mean(age)),
               Data_Table = dt[,.(avg = mean(age)), by = class],
               Tibble = ti %>% 
                       group_by(class) %>%
                       summarise(avg = mean(age)))
```

## Taking counts

Another function of interest is the '.N' function. This function will return the count of rows.

```{r counts}
microbenchmark(Data_Frame = df %>% 
                                 filter(credit_amount > 5000, age <35) %>%
                       nrow(),
               Data_Table = dt[credit_amount > 5000 & age < 35, .N ,],
               Tibble = ti %>% 
                       filter(credit_amount > 5000, age <35) %>%
                       nrow())
```

## Creating new columns

```{r new_col}
microbenchmark(Data_Frame = df$property <- paste(df$property_magnitude, df$housing),
               Data_Table = dt[,property := paste(property_magnitude, housing)],
               Tibble = ti$property <- paste(ti$property_magnitude, ti$housing))
```